


# Data Processing

RHoMIS data processing takes place in three stages.

1. Unit Extraction: We search through a survey for all of the units, crops, and livestock which were encountered. The user can examine these units and provide conversion factors so that everything is standardised. For crops and livestock, the user can check all of the entries and standardise (e.g. ensure that there are no duplicates with mispelling, such as "maize" and "maze").

2. Secondary Cleaning: Now that crop and livestock names are standardised (in step 1), we identify all of the calorie conversions and prices we need for a specific survey (e.g. based on crops identified in step one we need calorie and prices for "maize"). We use the units converted in step 1 to calculate crop and livestock prices (see [ Indicators Explained].

3. Indicator Calculation: Finally, with all conversions and units identified, we calculate all of the indicators possible for the survey. 


## Installation

Currently RHoMIS is not available on CRAN, but 
we hope one day it will be. In the meantime, to 
install the RHoMIS R-package, you will need to install `devtools`,
you can that by entering the following into the R console:

```{r eval=FALSE, echo=T}
install.packages("devtools")
```


Then to install the most recent version of the R-package you can enter 
the following command

```{r eval=FALSE, echo=T}
library(devtools)
devtools::install_github("https://github.com/RHoMIS/rhomis-R-package", force = TRUE)
```


## Initial Setup 

You should then ensure you have a folder containing the data
you would like to process. If using Rstudio interactively, set
your working directory to this folder using the `setwd()` command.
Your directory structure should look something like this:

```
📂 project
┃
┗📂 raw-data
  ┃
  ┗📜 raw-data.csv
```

In this case, your working directory should be the `project` folder.
If you do not have your own RHoMIS dataset, you can access one [here](https://raw.githubusercontent.com/RHoMIS/rhomis-R-package/4012fc3992ab9a1f8811d3e98a349d51aa66d581/inst/sample_local_project/raw-data/raw-data.csv) (you can save the file using `ctrl + s` or `cmd + s`)

## Stage 1: Unit Extraction

### Extracting Units

To extract new units from the data you are working with, run the
code below. Set the `dataFilePath` to the path of your RHoMIS 
raw-data file.

```{r eval=FALSE, echo=T}
library(rhomis)

extract_units_and_conversions_csv(
    file_path="raw-data/sample_project.csv",
    id_type = "string",
    proj_id = "test_prj",
    form_id = "test_frm")
```

If you now go back to your `project` directory, you should be able
to see that more files have been added:

```
📂 project
┃
┣📂 raw-data
┃ ┗📜 sample_project.csv
┃
┣📂 conversions_stage_1
┃ ┣📜 country.csv
┃ ┗📜 ...
┃
┗📂 .original_stage_1_conversions
  ┣📜 crop_yield_units.csv
  ┣📜 country.csv
  ┗📜 ...

```

The `conversions_stage_1` folder should will include the units which you will
verify and change.
The `.original_stage_1_conversions` folder includes an original copy of the units
pulled straight from the data. This folder will likely be hidden unless you 
change the settings on your file browser.


### Converting Units

__Please note, while you do not _have_ to convert units for the next step in
the calculations to run, uncoverted units will be ignored and will lead to 
unnecessary missing values__

In the `project/conversions_stage_1` directory, you will find a series of csv files 
with conversion factors that need to be filled in. Each file will look 
like the table below. The `survey_value` column indicates a value which was 
found in the survey. The `conversion` column shows the conversion factor 
which ought to be used. Where the conversion factor is unknown, the column
will display `NA`. Where the conversion factor is known it will be provided.
Known conversion factors are already embedded in the R-package.

```{r echo=F, eval=T}
knitr::kable(data.frame(list(
    unit_type = c("crop_unit", "crop_unit"),
    id_rhomis_dataset = c("x", "x"),
    survey_value = c("foo", "bar"),
    conversion = c(NA, NA)
)))
```

There are a range of conversions to enter and it is important to be aware
which units are needed. Each file is described in more detail below:


| File | Description | Example Survey Value | Example Conversion |
| :--- | :------- | :------- | :------- |
| honey_amount_to_l.csv | The conversion factor should be used to convert to kilograms/litres. | buckets_12_litre | 12 |
| country_to_iso2.csv | The conversion factor should be used to convert country names to [two-letter ISO country codes](https://www.iban.com/country-codes). This is important as it is used to find currency conversion factors  | uganda | UG |
| crop_name_to_std.csv | The name of crops entered in the survey. Often enumerators may specify "other" crops in a free-text-entry field. Sometimes these crops can be mispelled, or in a language other than English. Here we correct misspellings and translations into standard forms (__all lower case__)  |  maze | maize  |
| crop_price_to_lcu_per_kg.csv  |  The price units for crops which were sold. Please note that only one unit will be converted into a string, "total_income_per_year" will be converted to "total_income_per_year" as this is treated as a special case in the analysis scripts. | price_per_50kg_sack | 0.02 |
| crop_amount_to_kg.csv  |  The unit of crops which have been collected. This needs to be converted to kilograms | cart_250kg  | 250 |
| eggs_price_to_lcu_per_year.csv | The amount of money made per unit time for selling eggs. This needs to be converted into an income per year  | income for 3 months | 4 |
| eggs_amount_to_pieces_per_year.csv  | The number of eggs collected per year | pieces/day | 365 |
| fertiliser_amount_to_kg.csv|  The amount of fertiliser in kg |  sacks_25kg | 25  |
| livestock_name_to_std.csv  | The names of livestock entered in the survey. As with crop names, enumerators can also enter extra livestock names which are in a different language or mispelt. This conversion is used to standardise livestock names entered in the survey | catel | cattle |
| milk_price_to_lcu_per_l.csv  |  The amount of money made per unit volume or per unit time. For unit time, the options are "day", "week", "month", "year". These time units must be entered as strings. For the unit volume, numeric conversions must be entered| month | month |
|      |       |  0.3litre | 0.3  |
| milk_amount_to_l.csv  | The amount of milk  collected in litres per year. There are a number of exceptions which must be kept as text strings, and are dealt with in the analysis scripts (e.g. "per animal per day" and "l/animal/day" |  l/day  | 365 |
| land_area_to_ha.csv  | The amount of land in hectares |  acres  | 0.4  |


## Stage 2: Calculating Prices, Preparing Calorie Conversions, and Extracting Secondary Units

As with unit extraction, ensure you are in the project directory.
Then enter the commands below:

```{r eval=FALSE, echo=T}
library(rhomis)

calculate_prices_csv(
    file_path="raw-data/sample_project.csv",
    id_type = "string",
    proj_id = "my_project_id",
    form_id = "my_form_id"
)

```

You will now see that the directory structure again looks quite different:

```
📂 project
┃
┣📂 raw-data
┃
┃ ┗📜 sample_project.csv
┣📂 conversions_stage_1
┃ ┣📜 country.csv
┃ ┗📜 ...
┃
┗📂 .original_stage_1_conversions
┃ ┣📜 crop_yield_units.csv
┃ ┣📜 country.csv
┃ ┗📜 ...
┃
┣📂 conversions_stage_2
┃ ┣📜 crop_calories.csv
┃ ┣📜 mean_crop_pice_lcu_per_kg
┃ ┣📜 eggs_calories.csv
┃ ┣📜 mean_eggs_price_per_kg
┃ ┗📜 ...
┃
┗📂 .original_stage_2_conversions
  ┣📜 crop_calories.csv
  ┣📜 mean_crop_pice_lcu_per_kg
  ┣📜 eggs_calories.csv
  ┣📜 mean_eggs_price_per_kg
  ┗📜 ...



```

### Converting Calories, Prices, and Other Units

#### Calories and Prices

As with the units which were extracted, you will also need to verify
prices  and calorie values (in the `conversions_stage_2` folder). Prices will be in 
lcu/kg for crops, meat, and eggs (where lcu is local currency units).
Prices will be in lcu/l for milk and honey. And prices will be in 
lcu/animal for whole livestock sales. For calorie values, conversions will be in kcal/kg or kcal/l.

#### Secondary Units

Aside from prices and calories, there are other conversion tables to be verified, these units
depend on conversions from stage 1. For example `livestock_count_tlu` requires us to know
the standard livestock names (identified in stage 1). The secondary units currently 
processed in the R-package are:

| File | Description | Example Survey Value | Example Conversion |
| :--- | :------- | :------- | :------- |
| livestock_weight_kg.csv | The conversion factor used to estimate the amount of meat collected from a whole animal. | sheep | 25 |
| livestock_count_tlu.csv | "Tropical Livestock Units" (TLU) for individual animals. | cattle | 0.7 |

_Please note any folder with a dot before it will not be visible by default in most file explorers_

## Stage 3: Calculating Indicators

Using the calories and prices you have converted, you will then be able to produce
the final indicators. Either run the code below, or run the `03-calculate-final-indicators.R`
file. 


```{r eval=FALSE, echo=T}
library(rhomis)
calculate_indicators_local(
            file_path,
            id_type="string",
            proj_id,
            form_id,
            repeat_columns = pkg.env$repeat_columns,
            gender_categories = pkg.env$gender_categories)

```

You will now see the directory includes some
extra folders:

You will now see that the directory structure again looks quite different.
Extra folders and files have been created. There is a lot of information here,
for more explanation, please see [an explanation of indicators and outputs][Indicators Explained] 

```
📂 project
┃
┣📂 raw-data
┃ ┗📜 sample_project.csv
┃
┣📂 conversions_stage_1
┃ ┣📜 country.csv
┃ ┗📜 ...
┃
┗📂 .original_stage_1_conversions
┃ ┣📜 crop_yield_units.csv
┃ ┣📜 country.csv
┃ ┗📜 ...
┃
┣📂 conversions_stage_2
┃ ┣📜 crop_calories.csv
┃ ┣📜 mean_crop_pice_lcu_per_kg
┃ ┣📜 eggs_calories.csv
┃ ┣📜 mean_eggs_price_per_kg
┃ ┗📜 ...
┃
┣📂 .original_stage_2_conversions
┃ ┣📜 crop_calories.csv
┃ ┣📜 mean_crop_pice_lcu_per_kg
┃ ┣📜 eggs_calories.csv
┃ ┣📜 mean_eggs_price_per_kg
┃ ┗📜 ...
┃
┣📂 crop_data
┃ ┣📜 crop_consumed_kg_per_year.csv
┃ ┣📜 crop_harvest_kg_per_year.csv
┃ ┗📜 ...
┃
┣📂 indicator_data
┃ ┗📜 indicators.csv
┃
┣📂 livestock_data
┃ ┣📜 crop_consumed_kg_per_year.csv
┃ ┣📜 crop_consumed_kg_per_year.csv
┃ ┗📜 ...
┃
┣📂 off_farm_data
┃ ┣📜 offfarm_income_name.csv
┃ ┣📜 offfarm_who_control_revenue.csv
┃ ┗📜 ...
┃
┗📂 processed_data
┃  ┗📜 processed_data.csv
┃
┣📂 consumption_calorie_values
┃ ┣📜 crop_calories_consumed_kcal.csv
┃ ┣📜 milk_calories_consumed_kcal.csv
┃ ┗📜 ...
┃
┣📂 consumption_lcu_values
┃ ┣📜 value_crop_consumed_lcu.csv
┃ ┣📜 value_eggs_consumed_lcu.csv
┃ ┗📜 ...  
┃
┗📂 gender_control
┃
┣📂 gender_control_amounts_consumed
┃
┣📂 gender_control_income
┃
┗📂 ... 
```



