# Indicators Explained

The core RHoMIS survey contains over 750 variables.
These are used to calculate key indicators, such as 
`total_income_lcu_per_year`. 

This guide is designed to explain:

1. How some of the main indicators are calculated
2. How variables are linked
3. How you can explore these calculations and indicators yourself

In order to make the most of this guide, you should have a sample RHoMIS 
dataset, to see how to process a "Raw" RHoMIS dataset, and produce indicators
from this dataset, please see [data processing][Data Processing]

## Where to start

Firstly, we have created a list of indicators. To see
the names of all of the indicators, enter the following command:

```{r indicator_names, warning=FALSE, message=FALSE, eval=F}
names(indicator_list)
```

There are a lot of indicators here. Some of them you will find as standalone columns
in the RHoMIS Indicators Table. Others you will find in a looped format, in the processed
data. To find out more about an indicator, you can do as follows:

```{r individ_indicator_expl, warning=FALSE, message=FALSE, }
indicator_list[["hh_size_mae"]]
```
There is a lot of information here. So let's break it down. This is telling me
that the indicator is in a column format (in the indicator sheet). It requires a series
of individual columns, a series of looped columns and a conversion table. No other indicators
are needed to calculate "hh_size_mae". I can see it is calculated with the `calculate_household_size_members`. If I want to see where this function is stored in the 
R-package raw code I can enter the `?calculate_household_size_members` command into the console.
I can search for the tag `indicator_search_hh_size_mae` to see exactly where the indicator is 
calculated in this file.


Now lets look at a more complex indicator. `crop_income_lcu_per_year`:

```{r loop_indicator_expl, warning=FALSE, message=FALSE}
indicator_list[["crop_income_per_year"]]
```
I can see that this comes in a loop format. It relies two looped 
columns, a conversion table, and another indicator. Again I can find
the function that calculates the indicator, and use a search term to identify exactly 
the point where it is calcuated.

However, when one indicator depends on another, we might need do know
the nested relationship (i.e, what columns does the dependency rely on).
If we want to visualise it, we could do this:

```{r dependency_network_explained, warning=FALSE, message=FALSE}

plot_dependency_network(
    indicator_name="crop_income_per_year"
)


```
A lternatively,r if we want to extract the actual information, we can do this:

```{r depend_list,  warning=FALSE, message=FALSE}
find_nested_dependencies_list(
     indicator_name="crop_sold_kg_per_year",
     indicator_list=indicator_list,
     dependency_required="loop")

```
This would find all of the looped dependencies (recursively) for the 
indicator I am interested in.
